---
title: "heatmap"
author: "Elinor Sterner"
date: "2024-02-23"
output: html_document
keep_md: yes
---

```{r}
library(tidyverse)
library(reshape2)
setwd("/Users/elinorsterner/Documents/katzlab/allo/diff/making_heatmaps/")

og_tally <- read.csv("Subtrees_unaligned_meta_data.csv")

cluster_tally <- read.csv('Clusters_0.99_meta_data.csv')
```


Heat map dimensions:
X = life stages and uni- and multi-nucleate
Y = OG and Cluster
Z = proportion of cells


Subset clusters and clades sheets
```{r}

cluster_tally <- subset(cluster_tally, select = c(OG, cluster, porp_of_jv_in_cluster, porp_of_ad_in_cluster, porp_of_sz_in_cluster, porp_of_mu_in_cluster, porp_of_un_in_cluster, porp_of_nd_in_cluster))


og_tally <- subset(og_tally, select = c(OG, porp_of_jv_in_cluster, porp_of_ad_in_cluster, porp_of_sz_in_cluster, porp_of_mu_in_cluster, porp_of_un_in_cluster, porp_of_nd_in_cluster))
```

rename clusters and clades sheets
```{r}
renamed_cluster_tally <- cluster_tally %>% 
  rename(
    Juveniles = porp_of_jv_in_cluster,
    Adults = porp_of_ad_in_cluster,
    Schizonts = porp_of_sz_in_cluster,
    Multi = porp_of_mu_in_cluster,
    Uni = porp_of_un_in_cluster,
    NoNucData = porp_of_nd_in_cluster
    )

renamed_og_tally <- og_tally %>% 
  rename(
    Juveniles = porp_of_jv_in_cluster,
    Adults = porp_of_ad_in_cluster,
    Schizonts = porp_of_sz_in_cluster,
    Multi = porp_of_mu_in_cluster,
    Uni = porp_of_un_in_cluster,
    NoNucData = porp_of_nd_in_cluster
    )


```

Merge OG and cluster columns into one
  OG sheet skips this step
```{r}
renamed_cluster_tally <- renamed_cluster_tally|>
  mutate(OG_cluster = paste(OG, cluster, sep = '_'))

```


Flip to long format for heatmap
```{r}
long_cluster_tally <- pivot_longer(renamed_cluster_tally, 
                                cols = c(Juveniles, Adults, Schizonts, Multi, Uni, NoNucData),
                                names_to = "lifestage")

long_og_tally <- pivot_longer(renamed_og_tally, 
                                cols = c(Juveniles, Adults, Schizonts, Multi, Uni, NoNucData),
                                names_to = "lifestage")

```


Heat map by *cluster*
  
  Change the name of the output and the gradient breaks
```{r dev = c("png", "jpg")}

clusters <- ggplot(long_cluster_tally, aes(x = lifestage, y = OG_cluster, fill= value)) + 
  geom_tile() +
  theme(text = element_text(size=6)) +#need to adjust text size depending on n of rows
  scale_x_discrete(limits = c('Juveniles', 'Adults', 'Schizonts', 'Multi', 'Uni', 'NoNucData')) +
  ylab("Clustered Grabbed Clades") +
  scale_fill_stepsn(n.breaks = 5, colours = viridis::magma(5))
  # scale_fill_gradient2(low = "white", high = "deepskyblue4",
  #                      midpoint = 0.0,
  #                      breaks = seq(0, 1, 0.25), # adjust the breaks for the maximum proportion
  #                      limits = c(0, 1))# adjust the breaks for the maximum proportion


clusters

pdf('diff_clusters.pdf', width = 4 , height = 8) ## Change name and dimensions here

clusters

dev.off()# This gives an error if run alone, run the whole block


```

Heatmap by grabbed clade (by OG in your output)

  use the output sheet from countGeneFamily.py
  
```{r}
OGs_heat <- ggplot(long_og_tally, aes(x = lifestage, y = OG, fill= value)) + 
  geom_tile()+
   theme(text = element_text(size=8)) +
  scale_x_discrete(limits = c('Juveniles', 'Adults', 'Schizonts', 'Multi', 'Uni', 'NoNucData')) +
  ylab("Grabbed Clades") +
  scale_fill_stepsn(n.breaks = 5, colours = viridis::magma(5))
  # scale_fill_gradient2(low = "white", high = "deepskyblue4",
  #                      midpoint = 0.0,
  #                      breaks = seq(0, 1.2, 0.25), #adjust breaks
  #                      limits = c(0, 1.2))#adjust breaks

OGs_heat

pdf('diff_clades.pdf', width = 4 , height = 4) #change name

OGs_heat

dev.off()

```